---
title: "DLBCL_PG_inference"
author: "jenssle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r, warning=FALSE, message = FALSE}
library(reticulate)
library(basilisk)
library(RANN)
library(tidyverse)
library(cola)
library(psych)
library(readxl) 
library(broom)
library(scrime)
```

## Load data

First, load the 

```{r}
meta_ident_all <-
  readRDS(
    "~/Library/CloudStorage/OneDrive-JohannWolfgangGoetheUniversität/2021_DLBCL_proteogenomics/R_data/20220315_meta_identifier_cohort.rds"
  )
```

Load the latent space from the MuVi. The latent space comprises of latent factors from the integrated bulk RNA and global proteome expression data [MuVi model](https://github.com/MLO-lab/MuVI).

```{r}
Latent_space <-
  read.csv(
    "~/20220531_Z.csv" ,
    header = TRUE
  ) %>%
  rename(c("X" = "Prot_id" )) %>%
  column_to_rownames("Prot_id")

Latent_space_meta <-
  read.csv(
    "~/20220531_factor_meta.csv" ,
    header = TRUE
  ) %>%
  rename(c("X" = "factor")) %>%
  mutate(sig_at_0.1_either_joint = case_when(
    p_adj_pos_1_rna < 0.01 | p_adj_neg_1_rna < 0.01 ~ "True", #either pos or neg < 0.1 in RNA
    p_adj_pos_2_prot < 0.01 | p_adj_neg_2_prot < 0.01 ~ "True", #either pos or neg < 0.1 in Prot
    TRUE ~ "False" #means if none of the cases above is true
  ))
```

## Proteogenotype inference

To infer proteogenotypes, we  apply a graph-based community detection algorithm [Leiden algorithm](https://www.nature.com/articles/s41598-019-41695-z) in a consensus clustering framework. Therefore, we generate an adjacency matrix from the inferred and corrected latent factor space. With this adjacency matrix, Leiden-based communities are inferred iteratively. The resulting leiden-assignment x samples matrix is then subjected to a consensus clustering based approach, where there hamming-distance of the Leiden assignments is utilized for hierarchical clustering of the resulting proteogenotypes.

First, generation of the adjacency matrix.

```{r}
t_space <- Latent_space_filt %>% as.matrix() %>% t()

set.seed(42)
snn <- RANN::nn2(t(t_space), k=15)$nn.idx

snn 

adjacency_matrix <- matrix(0L, ncol(t_space), ncol(t_space))

rownames(adjacency_matrix) <- colnames(adjacency_matrix) <- colnames(t_space)
for(i in 1:ncol(t_space)) {
    adjacency_matrix[i,colnames(t_space)[snn[i,]]] <- 1L
}

#check that rows add to k
sum(adjacency_matrix[1,]) == 15
table(apply(adjacency_matrix, 1, sum))
```

Iterative Leiden-alg based community detection.

```{r}
#Retrieve names for labeling of the resulting leiden-communities x samples matrix
Latent_space_prots1 <- as.data.frame(Latent_space_filt) %>% tibble::rownames_to_column("Prot_id")

ident_global <- data.frame(Prot_id = Latent_space_prots1$Prot_id)

for(i in 1:1000) {

set.seed(sample(1:1000,1))
partition_new <- leiden(adjacency_matrix, 
                        partition_type = c("RBConfigurationVertexPartition"), # implements modularity to derive partition
                        resolution_parameter = 1, 
                        # weights = weights_obj,
                        n_iterations = -1L, # set to negative --> algorithm runs until partition is not improved
                        seed = sample(1:1000,1)
                        )

Latent_space_prots1 <- as.data.frame(Latent_space_filt) %>% tibble::rownames_to_column("Prot_id")

ident_partition <- cbind(Latent_space_prots1$Prot_id, partition_new) %>% as.data.frame()

colnames(ident_partition) <- c("Prot_id", i)

ident_global <- left_join(ident_global, ident_partition, by = "Prot_id")

}

#get rownames and transpose the resulting df
ident_global <- ident_global %>%
  column_to_rownames("Prot_id") %>%
  t() %>%
  as.data.frame() 

#mutate all variables to numeric
ident_global_m <- mutate_all(ident_global, function(x) as.numeric(as.character(x)))
```

### Consensus clustering of iterative leiden assignments

To run the consensus clustering, we rely on the package [cola](https://www.bioconductor.org/packages/release/bioc/html/cola.html). First, register the "take all" method for top values.

```{r}
register_top_value_methods(
 
  all = function(mat) {
    
    all <- c(runif(n = nrow(mat), min = 1, max = 1))
 
  }
 
)
```

Second, register hclust on the basis of hamming distance.

```{r}
register_partition_methods(
    hclust_hamming = function(mat, k) {
      
      library(tidyverse)
      library(e1071)
      
      mat_fac <- mutate_all(as.data.frame(mat),as.factor)
        
      
      ham_dist_fac <- e1071::hamming.distance(as.matrix(t(mat_fac)))
      
      res_clust <- cutree(hclust(as.dist(ham_dist_fac)), k)
      
      res_clust
    }
)
```

Run the consensus clustering using the hamming distance and hierarchical clustering

```{r}
leiden_rl <- run_all_consensus_partition_methods(
  as.matrix(ident_global_m),
  top_value_method = c("all"),
  partition_method = c("hclust_hamming"),
  mc.cores = 35,
  top_n = nrow(ident_global_m),
  max_k = max(ident_global_m),
  scale_rows = FALSE
  )
```

Now search for the best partitions of the repetitive Leiden clusters.

```{r}
suggest_best_k(leiden_rl)
```

Extract the results.

```{r}
res_leiden = leiden_rl["all", "hclust_hamming"]
```

Plot the respective consensus heatmap.

```{r}
consensus_heatmap(res_leiden, 7)
```

Highlight the membership heatmap.

```{r}
membership_heatmap(res_leiden, k = 7)
```

Show discrimination by dimension reduction.

```{r}
dimension_reduction(res_leiden, k = 7, method = "UMAP")
```

Now, extract the respective information of the class membership.

```{r}
res_class <- get_classes(res_leiden, k=7) %>%
  rownames_to_column("Prot_id")

head(res_class)
```