---
title: "scRNA_ATAC_integration"
author: "jenssle"
date: "`r Sys.Date()`"
output: html_document
---

### Load packages 

```{r}
library(ArchR)
library(tidyverse)
library(Seurat)
library(SeuratObject)
library(scater)
library(scran)
library(scuttle)
library(patchwork)
library(bluster)
library(batchelor)
library(miloR)
library(scCustomize)
```

### Load seurat object 

The scRNA data and the scATAC data were processed independently and finally merged for all cells with present meta data and sufficient quality metrics in scRNA and scATAC.

```{r}
seurat_ob <- readRDS("seurat_ob_RNA_ATAC.rds")
```

### WNN integration

The weighted nearest neighbor (WNN) procedure implemented in Seurat v4 is designed to integrate multiple types of data that are collected in the same cells to define a single unified representation of single-cell multimodal data. For each cell, the procedure learns a set of modality weights, which reflect the relative information content for each data type in that cell. 

(1) Constructing independent k-nearest neighbor (KNN) graphs for each modalities. 
(2) Performing within and across-modality prediction 
(3) Calculating cell-specific modality weights. 
(4) Calculating a WNN graph.

Using the implemented weighted-nearest-neighbour integration method, we integrate the reduced dimension from RNA (pca) amd the harmonized LSI from ATAC data.

```{r}
seurat_ob <- FindMultiModalNeighbors(seurat_ob, reduction.list = list("pca", "harmony_lsi"), dims.list = list(1:50, 2:30))
```

Run UMAP on the dimred.

```{r}
seurat_ob <- RunUMAP(seurat_ob, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")
```

Plot for RNA integrated PCA. 

```{r}
DimPlot(seurat_ob, group.by = c("celltype", "orig.ident"), reduction = "umap", label = FALSE, label.size = 2.5, repel = TRUE, raster = FALSE)
```

Plot for ATAC harmony LSI UMAP. 

```{r}
seurat_ob <- RunUMAP(seurat_ob, reduction = 'harmony_lsi', dims = 2:30, reduction.name = "umap.atac", reduction.key = "atacUMAP_")

DimPlot(seurat_ob, group.by = c("celltype", "orig.ident"), reduction = "umap.atac", label = FALSE, label.size = 2.5, repel = TRUE,  raster = FALSE)
```

Plot for the integrated dimension.

```{r}
DimPlot(seurat_ob, group.by = c("celltype", "orig.ident"), reduction = "wnn.umap", label = FALSE, label.size = 2.5, repel = TRUE, raster = FALSE)
```

### Milo DA on WNN

First, transform the object into sce. 

```{r}
seurat_ob <- RunPCA(seurat_ob, nn.name = "weighted.nn", reduction.name = "wnn.pca", reduction.key = "wnnPCA_")

DefaultAssay(seurat_ob) <- "RNA"

sce.ob <- as.SingleCellExperiment(seurat_ob)

DefaultAssay(seurat_ob) <- "integrated"
#map also the already generated umap to the sce object
reducedDim(sce.ob, "wnn.umap") <- Embeddings(seurat_ob, reduction = "wnn.umap")

reducedDim(sce.ob, "wnn.pca") <- Embeddings(seurat_ob, reduction = "wnn.pca")

milo.sce<- Milo(sce.ob)
```

Add the precomputed WNN graph from the seurat processing. 

```{r}
set.seed(42)
milo_graph <- buildFromAdjacency(seurat_ob@graphs$wknn, k = 20, is.binary = TRUE) 

graph(milo.sce) <- miloR::graph(milo_graph)
```

Define the nhoods. 

```{r}
set.seed(42)
milo.sce <- makeNhoods(milo.sce, prop = 0.05, k = 200, refinement_scheme="graph", refined = TRUE)

plotNhoodSizeHist(milo.sce)
```

Counting the cells. 

```{r}
set.seed(42)
milo.sce <- countCells(milo.sce, 
                                    meta.data= as.data.frame(colData(milo.sce)),
                                    sample = "orig.ident")

head(nhoodCounts(milo.sce))
```

Generate the design matrix. 

```{r}
milo_design <- data.frame(colData(milo.sce))[,c("orig.ident", "celltype")]

pg4_vec <- c("cornell_105", "cornell_110", "cornell_122", "cornell_130", "cornell_137", "cornell_19", "cornell_2", "cornell_21")

milo_design <- milo_design %>%
  mutate(pg4 = case_when(orig.ident %in% pg4_vec ~ "pg4", TRUE ~ "other")) %>%
  distinct(orig.ident, .keep_all = TRUE)

rownames(milo_design) <- milo_design$orig.ident

milo_design
```

Compute the neighbourhood connectivity. This function will calculate Euclidean distances between single-cells in a neighbourhood using the same dimensionality as was used to construct the graph. Since we do not have a dimreduction but started directly with the multi-modal graph, we compute the PCA on the wnn graph and supply this as the dimred to the function.

```{r}
set.seed(42)
milo.sce <- calcNhoodDistance(milo.sce, d=20, reduced.dim = "wnn.pca")
```

Run the testing. 

```{r}
da_results <- testNhoods(milo.sce, 
                         design = ~ pg4, 
                         design.df = milo_design,
                         reduced.dim = "wnn.pca",
                         fdr.weighting="graph-overlap"
                         )
```

Visualize the readout. 

```{r, fig.height=6, fig.width=12}
library(scales)
set.seed(42)
milo.sce <- buildNhoodGraph(milo.sce)

## Plot single-cell UMAP
pl <- plotReducedDim(milo.sce, dimred = "wnn.umap", colour_by="celltype", 
                          text_size = 3, point_size=0.05) +
  guides(fill="none") +
  scale_color_manual(values = hue_pal()(13))

## Plot neighbourhood graph
nh_pl <- plotNhoodGraphDA(milo.sce, da_results, layout="wnn.umap",alpha=0.1)
  
pl + nh_pl +
  plot_layout(guides="collect")
```